---
title: "Fake Test "
author: "David John Baker"
date: "02/03/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Question One 

```{r}
library(tidyverse)

at_health_facilities <- read_csv("data/at_health_facilities.csv")

at_health_facilities

```

* How Many Countries are reported in the dataset?

```{r}

at_health_facilities %>%
  select(iso3) %>%
  distinct() %>%
  summarise(n = n())

```

* What is the difference between the minimum and maximum year with valid data for each country?

```{r}
# 

at_health_facilities %>%
  rename(age_15_17 = `age 15-17`,
         age_20_34 = `age 20-34`) %>%
  mutate(age_20_34 = as.double(age_20_34)) %>%
  filter(age_15_17 != is.na(age_15_17) | age_20_34 != is.na(age_20_34)) %>%
  group_by(iso3) %>%
  summarise(max_year = max(year), min_year = min(year)) %>%
  mutate(year_diff = max_year - min_year) %>%
  arrange(-year_diff)

```

* How many countries reported data in 3 or more years? 

```{r}
at_health_facilities %>%
  rename(age_15_17 = `age 15-17`,
         age_20_34 = `age 20-34`) %>%
  mutate(age_20_34 = as.double(age_20_34)) %>%
  filter(age_15_17 != is.na(age_15_17) | age_20_34 != is.na(age_20_34)) %>%
  group_by(iso3) %>%
  summarise(number_iso3_entries = n()) %>%
  filter(number_iso3_entries >= 3) 

```
 
* Which countries reported 100% incidence for at least one year in either age group?
 
```{r}
at_health_facilities %>%
  rename(age_15_17 = `age 15-17`,
         age_20_34 = `age 20-34`) %>%
  mutate(age_20_34 = as.double(age_20_34)) %>%
  filter(age_15_17 != is.na(age_15_17) | age_20_34 != is.na(age_20_34)) %>%
  filter(age_15_17 == 100 | age_20_34 == 100) %>%
  select(iso3) %>%
  distinct()

```

## Question Three

> Describe what columns a tidy layout for this data would have and why.

In order for this data to be tidy, each variable needs to form a column and each separate observation as a row.
In this setting, each column does have some variable information there, but there are two pieces of info.
The first is the year, the second is if it is an estimation, high, or low value.
Further, the `>- 95%` has meaning in addition to just the number itself and should be accounted for. 

> Write a function that takes the name of a file containing this table as input and returns a tidy version of the table.
The function should replace all - and >95% values with NA.
The body of the function may contain one or more pipelines and may create temporary or intermediate variables, but may not contain any loops.

```{r}
hiv <- read_csv("data/infant_hiv.csv")

# First melt entire dataframe to get columns and values together, then separate out the column info. 

hiv %>%
  pivot_longer(-ISO3, names_to = "year", values_to = "percent") %>%
  separate(year, c("year","label")) %>%
  mutate(year = as.integer(year)) %>% 
  mutate(percent = str_remove_all(string = percent, pattern = "-")) %>%
  mutate(percent = str_remove_all(string = percent, pattern = ">95%")) %>%
  mutate(percent = str_remove_all(string = percent, pattern = "%")) %>%
  mutate(percent = as.integer(percent)) 


clean_hiv_table <- function(hiv_table) {
  hiv_table %>%
    pivot_longer(-ISO3, names_to = "year", values_to = "percent") %>%
    separate(year, c("year", "label")) %>%
    mutate(year = as.integer(year)) %>%
    mutate(percent = str_remove_all(string = percent, pattern = "-")) %>%
    mutate(percent = str_remove_all(string = percent, pattern = ">95%")) %>%
    mutate(percent = str_remove_all(string = percent, pattern = "%")) %>%
    mutate(percent = as.integer(percent))
  
}

clean_hiv_table(hiv)

```

That took ten minues ! ]


## Plot Re-Creation

The file ranking.csv contains two columns:

The ID of an item being rated.
A rating, which is one of “negative”, “positive”, “indifferent”, or “wtf” (meaning the respondent didn’t understand the question).
There are multiple ratings for each item.

The plot below shows this data:

Each dot represents one item i.
The size of the circles shows the total number of ratings for item i.
The X coordinate for item i is the percentage of ratings for that item that are “negative”.
The Y coordinate for item i is the percentage of ratings for that item that are “positive”.
The regression line is created using the ‘lm’ method.

```{r}
ranking <- read_csv("data/ranking.csv")

ranking %>%
  group_by(item)

# 284 total items 

ranking %>%
  count(item) %>%
  ggplot(aes(x = reorder(item,n), y = n)) +
  geom_bar(stat = "identity") + 
  coord_flip()


ranking %>%
  count(item) -> count_table 

ranking %>%
  group_by(item, rank) %>%
  summarise(counts = n()) %>%
  mutate(per_item_total = sum(counts)) %>%
  mutate(relative_freq = counts/per_item_total) %>%
  filter(rank == "positive" | rank == "negative") %>%
  select(item, rank, relative_freq) %>% 
  pivot_wider(names_from = rank, values_from = relative_freq) %>%
  left_join(count_table, by = "item") %>%
  ggplot(aes(y = positive, x = negative, size = n, alpha = n)) +
  geom_point() +
  geom_smooth(method = "lm") 
  

```


