---
title: "Practice 2 -- Faster"
author: "David John Baker"
date: "30/03/2020"
output: html_document
---


The file `at_health_facilities.csv` contains a tidy dataset with four columns: 

- The ISO3 code of the country that reported data. 
- The year for which data was reported. 
- The percentage of HIV-positive children born to HIV-positive mothers age 15–17. 
- The percentage of HIV-positive children born to HIV-positive mothers age 20–34.

Please answer the following questions:

* How many countries reported data?
* What is the difference between the minimum and maximum year with valid data for each country?
* How many countries reported data in 3 or more years?
* Which countries reported 100% incidence for at least one year in either age group?


```{r}
library(tidyverse)

health <- read_csv("david_practices/data/at_health_facilities.csv")

# How many countries reported data?
health %>%
  select(iso3) %>%
  distinct() %>%
  nrow()

# What is the difference between the minimum and maximum year with valid data for each country?

health %>%
  mutate(`age 20-34` = na_if(`age 20-34`, "-")) %>%
  mutate(`age 20-34` = as.double(`age 20-34`)) %>%
  filter(!is.na(`age 20-34`)) %>%
  group_by(iso3) %>%
  mutate(max_year = max(year),
         min_year = min(year)) %>%
  mutate(diff = min_year - max_year) %>%
  select(iso3, diff, `age 15-17`, `age 20-34`) %>%
  distinct()
  
# How many countries reported data in 3 or more years?

health %>%
  group_by(iso3) %>%
  select(iso3, year) %>%
  summarise(n = n()) %>%
  filter(n >= 3) %>%
  nrow()

# Which countries reported 100% incidence for at least one year in either age group?

health %>%
  mutate(`age 20-34` = na_if(`age 20-34`, "-")) %>%
  mutate(`age 20-34` = as.double(`age 20-34`)) %>%
  filter(`age 15-17` == 100 | `age 20-34` == 100) %>%
  distinct(iso3) 

# 20 Minutes! 
```

## Question 2

* Done! 20 min done !

## Question 3

* Use `infant_hiv.csv` to tidy the data 

Describe what columns a tidy layout for this data would have and why.
Write a function that takes the name of a file containing this table as input and returns a tidy version of the table.
The function should replace all - and >95% values with NA.
The body of the function may contain one or more pipelines and may create temporary or intermediate variables, but may not contain any loops.

1. Each column with the name like YEAR rating, needs be split up so there is both a year as a grouping variable as well as the rating next to it. Then each row will be tidy!

* function takes name of file and spits out table 

* [ ] replace all - and >95 with na
* [ ] No loops, only pipes 

```{r}
infant <- read_csv("david_practices/data/infant_hiv.csv")

clean_infant_hiv_table <- function(x){
  if(!is.character(x)){
    print("You need to pass the file name in as a string!")
  }
  infant <- read_csv(x)
  infant %>%
  pivot_longer(cols = `2009 est`:`2017 lo`, names_to = "year_rating", values_to = "rates") %>%
  separate(col = year_rating, into = c("year","est")) %>%
  mutate(rates = na_if(x = rates, "-")) %>%
  mutate(rates = na_if(x = rates, ">95%")) %>%
  mutate(rates = str_remove_all(string = rates, pattern = "%")) %>%
  mutate(rates = as.integer(rates))
}


clean_infant_hiv_table("david_practices/data/infant_hiv.csv")  

```
# 45 minutes left 

## Recreate this plot 

```{r}
ranking <- read_csv("david_practices/data/ranking.csv")

ranking %>%
  group_by(item) %>%
  add_count(name = "ratings_for_item") %>%
  ungroup(item) %>%
  group_by(item, rank) %>%
  add_count(name = "per_level") %>%
  distinct() %>%
  arrange(item, rank) %>%
  mutate(total_ratio = per_level / ratings_for_item) %>%
  filter(rank == "positive" | rank == "negative") %>% 
  pivot_wider(names_from = rank, values_from = total_ratio) %>%
  select(item, positive) -> positive_table 
  
ranking %>%
  group_by(item) %>%
  add_count(name = "ratings_for_item") %>%
  ungroup(item) %>%
  group_by(item, rank) %>%
  add_count(name = "per_level") %>%
  distinct() %>%
  arrange(item, rank) %>%
  mutate(total_ratio = per_level / ratings_for_item) %>%
  filter(rank == "positive" | rank == "negative") %>% 
  pivot_wider(names_from = rank, values_from = total_ratio) %>%
  select(item, negative) -> negative_table   

positive_table %>%
  left_join(negative_table,by = "item")
  
  
  ggplot(aes(x = negative, y = positive, size = ratings_for_item)) +
  geom_point()
  

```





