---
title: "Learn to Pivot!"
author: "David John Baker"
date: "21/03/2020"
output: html_document
---

Let's start by importing what we need for this lesson!

```{r, echo=TRUE, results='hide'}
# Import Needed Data and Libraries
library(tidyverse)
library(knitr)
study_data <- read_csv("../data/study_data.csv")
study_data_long <- read_csv("../data/study_data_long.csv")
student_data <- read_csv("../data/student_data.csv")
store_data <- read_csv("../data/store_data.csv")
```

Here's the data we're going to be working with.

```{r}
study_data
```

The same data, but in a different shape.

```{r}
study_data_long
```

Let's now look at the code needed to change data from wide to long.
In order to do this, you will need to tell `pivot_longer()` *four* different arguments. 

* data: what data are you working with?
* cols: what are the columns that look like variables, but are really values?
* names_to: What kind of data is living in those column headers?
* values_to: What kind of values are living in those columns? I also like to think about this as what unit is the variable measured on? 

```{r,eval=FALSE}
study_data

# pivot_longer(data = ______ , cols = _______, names_to = ______, values_to = ______)
```

As a group!

Take a look at this new data and see if you can fill in the blanks in order to get this new data from wide format to long.
In this data set we have data from a few different students and their daily grades from one week of the year.
We want every row to contain the name of the student, the day of the week, and their grade.

> Use the example above as a model if you get stuck. 

Remember to first uncomment the line so you can run it! 


```{r}
student_data

# pivot_longer(data = ______ , cols = _______, names_to = ______, values_to = ______)
```

To end, let's try one more, but this time I am only going to give you the data and tell you that I want you to take this data and make it longer!
The data here contains sales of products from a family owned store over the past few months.
Here we want to take the sales from the different months and make it so that we have a variable called `month` and a variable called number of `sales`. 
There are a few other variables in there too, but just focus on the ones you want to pivot and `pivot_longer()` will take care of the rest. 

```{r}
store_data

# Your Code Here! 

```

We covered a lot of ground in the last 15 minutes.
We reviewed the idea of tidy data and learned two new terms to describe the shape of data.
What are they?

> ??? 

We also looked at the first of two `pivot` functions, `pivot_longer()` which takes wide data and makes it long.
The function `pivot_longer()` needs four arguments at minimum.
What are they?

> ???

Let's take a quick break, then do `pivot_wider()`...


