---
title: "Learn to Pivot!"
author: "David John Baker"
date: "20/03/2020"
output: html_document
---

!! Change Learner Persona to Undergrad 

Today we continue learning about the idea of [tidy data](https://r4ds.had.co.nz/tidy-data.html).
We'll build on some of concepts from last time so you can start to manipulate some of your own datasets and change the way they are organised with only a few lines of code.

## The Shape of Data 

How many people here have often heard that your brain is a computer?

* Engages people

This is partially true based on how you define computation and how far you are willing to stretch the metaphor, but even though people say your brain _computes_ things, human brains do not work exactly like the computer that's sitting in front of you. 

If your brain and computers worked exactly the same, you'd imagine that both you and your computer would always want things in the same format to make it easier for both parties involved.
As psychologists we know this isnt't always the case.

In practice (and reality), humans and computers often need to be flexible in how they think about things.
What works well for one situation for a brain or a computer might not work well in another.

We see this all the time in reserach.

Take a look at the following table of data.

* Example of data in wide format 

A scientist is running a study where they need to collect data at multiple timepoints on a smoking cessation study and need to track number of cigarrets smoked and need to enter those values in by hand or maybe a teacher needs to record daily grades, this is a way that is familiar to many people who work with spreadsheets.

The problem is that when the data is structured like this, in order to do some sort of calculation like count the total number of cigarettes smoked or the average grade per student, most of the software we have seen thus far needs the data to operate over columns as opposed to rows. 

For example, look at this new data frame.
It contains the exact same data as the table before, but has a different shape to it.

* Long Data Frame

Both shapes of data are tidy but one is probably easier to enter data into and the other is in a format that works well with a lot of software that makes our jobs easier.
Often we need to change our data from one of these shapes to the other.
And imporantly, each of these shapes of data has a specific name.

When some of the columns of a dataset have certain values, rather than the explicit name of a variable, the data is called *wide data*.
When an observation is scattered across many rows, the data is called *long data*. 

Take a second to talk with your partner and answer this question:

Given the dataset shown here, which of the following choices best describes the current state of the data using terminology we just learned.

- A. The dataset is wide and tidy
- B. The dataset is long and tidy
- C. The dataset is wide and not tidy
- D. The dataset is long and not tidy.

## Pivoting 

Great, now that we're on the same page, let's learn how to change between the two.
When you want to change the shape of data in the way that we just talked about in the tidyverse, you use the `pivot` functions.
If you have wide data and want to make it long, you use `pivot_longer()`; if you have long data and want to make it wide, you use `pivot_wider()`.
We're going to first look at `pivot_longer()` then in another lesson can explore `pivot_wider()`.

If you are going to use `pivot_longer()`, we can assume that your data is in a wide format.
Since your data is in a wide format we know that some of the column names are actually _values_ that look like variables.
There can be a lot of terms to keep track of here so let's start one by one.

In general, the first argument for any `tidyverse` function is going to be the data we want to manipulate.
Let's start with that:

```{r}
library(tidyverse)

df 

#pivot_longer(data = ,cols = ,names_to = ,values_to = )
```


In the example below from before, which of our columns here look like variables, but are really values?

* Wide data from before

We have now identified our columns that are hiding as variables but are really values.
Identifying columns that look like variables, but are really values is the first step in moving data from wide to format.
If you remember this, you will remember that this is the second argument that the `pivot_longer()`  function takes `id_cols =`.
The function `pivot_longer()` makes it easy for us and allows for us to use `select()` like notation to select out what columns we have identified.

If you need a refresher on all the things `select()` can do, check that out [here]().

Great, now that we have our data and identifed what columns are values hiding as variables, we can now tell `pivot_longer()` what we have found.
We now need to tell `pivot_longer()` both what kind of data was in our column names as well as what kind of data is in the column with the value as a header.
In our case here, the _____________.

Let's fill that in with our example.

```{r}

```


We now have everything we need to move from wide data to long.
Let's run this to see if it worked.

```{r}

```

Great! 
We've done it.

Let's try another one as a group, but not full on.
Take a look at this new data and see if you can fill in the blanks in order to get this new data from long format to wide.


```{r}
# pivot_longer(df, cols = , names_to = , values_to = )

```

Again, nice work.

To end, let's try one more, but this time I am only going to give you the dataframe and tell you that I want you to take this dataframe and make it longer!

```{r}
# df_3
```


Great.
We covered a lot of ground in the last 15 minutes.
We reviewed the idea of tidy data and learned two new terms to describe the shape of data.
What are they?

We also looked at the first of two `pivot` functions, `pivot_longer()` which takes wide data and makes it long.
The function `pivot_longer()` needs four arguments at miminum.
What are they?

Great, let's take a quick break, then do `pivot_wider()`.

